\vspace{1.0cm} \LARGE \bf Revealing the competitive dynamics of influenza viruses through evolutionary cartography \\ or \\ Charting antigenic drift and evolutionary success of seasonal influenza lineages

### 2013-01-22 flux.tex
BEAST 1.5 will not supported most of the analysis in the paper; do not cite.

Here, the phylogenetic tree $\tree$ is estimated using sequence data for viruses $1,\ldots,\vn$ according to well established methods implemented in the software package BEAST \cite{BEAST, BEAST17}.

@@

Top-level priors for $1/\mdssd^2$, $\beta$, $1/\virussd^2$, and $1/\serumsd^2$ are assumed to follow diffuse $\mbox{Gamma}(a, b)$ distributions  with $a=0.001$ and $b=0.001$.
Under the full model, the posterior probability of observing virus and serum locations given immunological data is factored
\begin{equation}
	p(\viruses,\sera | \mathbf{H}) \propto p(\mathbf{H} | \viruses, \sera, \ses, \ves, \mdssd) \; 
	p(\viruses | \beta, \virussd, \tree) \;
	p(\sera | \beta, \serumsd) \; 
	p(\ses, \ves, \mdssd, \beta, \virussd, \serumsd, \tree).
\end{equation}
We sample from this posterior distribution using the Markov chain Monte Carlo (MCMC) procedures implemented in the software package BEAST \cite{BEAST,BEAST17}.
Metropolis-Hastings proposals include transition kernels that translate individual virus and serum locations $\virus_i$ and $\serum_j$ and individual virus effects $\ve_i$ and serum effects $\se_j$, and other transition kernels that scale the entire set of virus and serum locations $\viruses$ and $\sera$ and that scale parameters $\mdssd$, $\beta$, $\virussd$ and $\serumsd$.
For the present analysis, a two-step approach was taken to sample phylogenies, where a posterior sample of phylogenies was gathered using sequence data and then, in the cartographic analysis, trees from this set were randomly proposed and accepted following the Metropolis-Hastings algorithm \cite{Pagel04}.

@@

Phylogenetic trees were sampled for A/H3N2, A/H1N1, B/Vic and B/Yam using BEAST \cite{BEAST} and incorporated the SRD06 nucleotide substitution model \cite{Shapiro06}, a coalescent demographic model with constant effective population size and a strict molecular clock across branches.
MCMC was run for 60 million steps and trees were sampled every 50,000 steps after allowing a burn-in of 10 million steps, yielding a total sample of 2000 trees.
These trees were treated as a discrete set of possibilities when subsequently sampled in the BMDS analysis \cite{Pagel04}.

@@

Source code implementing the cartographic models has been made fully available as part of the software package BEAST \cite{BEAST, BEAST17}, and can be downloaded from its Google code repository (http://code.google.com/p/beast-mcmc/).
Antigenic, genetic and surveillance data is available as supplemental information \tbc{PREPARE}.

>>>>

Here, the phylogenetic tree $\tree$ is estimated using sequence data for viruses $1,\ldots,\vn$ according to well established methods implemented in the software package BEAST \cite{BEAST17}.

@@

Top-level priors for $1/\mdssd^2$, $\beta$, $1/\virussd^2$, and $1/\serumsd^2$ are assumed to follow diffuse $\mbox{Gamma}(a, b)$ distributions  with $a=0.001$ and $b=0.001$.
Under the full model, the posterior probability of observing virus and serum locations given immunological data is factored
\begin{equation}
	p(\viruses,\sera | \mathbf{H}) \propto p(\mathbf{H} | \viruses, \sera, \ses, \ves, \mdssd) \; 
	p(\viruses | \beta, \virussd, \tree) \;
	p(\sera | \beta, \serumsd) \; 
	p(\ses, \ves, \mdssd, \beta, \virussd, \serumsd, \tree).
\end{equation}
We sample from this posterior distribution using the Markov chain Monte Carlo (MCMC) procedures implemented in the software package BEAST \cite{BEAST17}.
Metropolis-Hastings proposals include transition kernels that translate individual virus and serum locations $\virus_i$ and $\serum_j$ and individual virus effects $\ve_i$ and serum effects $\se_j$, and other transition kernels that scale the entire set of virus and serum locations $\viruses$ and $\sera$ and that scale parameters $\mdssd$, $\beta$, $\virussd$ and $\serumsd$.
For the present analysis, a two-step approach was taken to sample phylogenies, where a posterior sample of phylogenies was gathered using sequence data and then, in the cartographic analysis, trees from this set were randomly proposed and accepted following the Metropolis-Hastings algorithm \cite{Pagel04}.

@@

Phylogenetic trees were sampled for A/H3N2, A/H1N1, B/Vic and B/Yam using BEAST \cite{BEAST17} and incorporated the SRD06 nucleotide substitution model \cite{Shapiro06}, a coalescent demographic model with constant effective population size and a strict molecular clock across branches.
MCMC was run for 60 million steps and trees were sampled every 50,000 steps after allowing a burn-in of 10 million steps, yielding a total sample of 2000 trees.
These trees were treated as a discrete set of possibilities when subsequently sampled in the BMDS analysis \cite{Pagel04}.

@@

Source code implementing the cartographic models has been made fully available as part of the software package BEAST \cite{BEAST17}, and can be downloaded from its Google code repository (http://code.google.com/p/beast-mcmc/).
Antigenic, genetic and surveillance data is available as supplemental information \tbc{PREPARE}.
